Revisiting the Austral Spring Extratropical Southern Hemisphere zonally asymmetric circulation using complex Empirical Orthogonal Functions

### 2022-04-20 analysis/paper/shceof.Rmd
Quita latex2expr

@@

(ref:rotations-cap) a) Correlation of the Real and Imaginary cEOF1 with wave 1 amplitude of Total Ozone Column averaged between `r knitr::combine_words(LatLabel(o3wave_lats))` and b) $R^2$ between the Real and Imaginary cEOF2 and the ONI index for different values of rotation. The vertical line marks the selected rotation, which maximises de correlation.

```{r rotations, fig.cap="(ref:rotations-cap)", fig.height=2.5, fig.width=width_column, include=FALSE}
r2 <- expression(`$r^2$` = paste("", "r", phantom()^{
  paste("2")
}, "", ""))

rotations_o3 %>%
  ggplot(aes(rotation, cor)) +
  geom_vline(xintercept = best_rotation_cEOF1) +
  geom_hline(yintercept = 0,  size = 0.2, color = "gray50") +
  geom_line(aes(color = part)) +
  scale_x_continuous("Rotation", breaks = seq(-pi, pi, by = 90*pi/180),
                     labels = function(x) paste0(round(x*180/pi, 5), "º")) +
  scale_y_continuous("Correlation") +
  scale_color_brewer(NULL, palette = "Dark2", labels = lab_cplx) +

  rotations_cEOF2 %>%
  melt(id.vars = "rotation") %>%
  ggplot(aes(rotation, value^2)) +
  geom_hline(yintercept = 0,  size = 0.2, color = "gray50") +
  geom_vline(xintercept = best_rotation_cEOF2) +
  geom_line(aes(color = variable)) +
  scale_x_continuous("Rotation", breaks = seq(-pi, pi, by = 90*pi/180),
                     labels = function(x) paste0(round(x*180/pi, 5), "º")) +
  scale_y_continuous(r2) +
  scale_color_brewer(NULL, palette = "Dark2", labels = lab_cplx) +

  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = "a", tag_suffix = ")")

@@

mgp <- latex2exp::TeX('Geopotential height ($m^2s^{-1}$)')
```

@@

```{r eof1-regr-gh, fig.cap = "(ref:eof1-regr-gh-cap)", fig.height=1.2*width_column, fig.width=width_column, dependson="plot_regr"}
plot_regr(regr , "cEOF1", c(200, 50), "hgt") +
  scale_fill_divergent(mgp,
                       breaks = AnchorBreaks(0, exclude = 0),
                       guide = guide_colorsteps_bottom(15, even.steps = TRUE), 
                       limits = c(-2000, 2000)) 
```

@@

```{r eof2-regr-gh, fig.cap = "(ref:eof2-regr-gh-cap)", fig.height=1.3*width_column, fig.width=width_column, dependson="plot_regr"}
plot_regr(regr, "cEOF2", c(200, 50), "hgt") +
  scale_fill_divergent(mgp, 
                       breaks = AnchorBreaks(0, 200, exclude = 0),
                       guide = guide_colorsteps_bottom(15, even.steps = TRUE), 
                       limits = c(-1000, 1000))
```

@@

temp %>% 
  ggplot(aes(Mod(hgt), amplitude)) +
  scale_x_continuous("EOF1 Amplitude") +
  scale_y_continuous("O3 Wave-1\nAmplitude") +

  temp %>% 
  copy() %>%
  na.omit() %>% 
  qwrap(phase = c(0, 2*pi) ~ c(-pi, pi)) %>% 
  ggplot(aes(Arg(hgt), phase)) +
  scale_x_continuous("EOF1 Phase",
                     breaks = seq(-pi, pi, by = 15*pi/180),
                     labels = function(x) paste0(round(x*180/pi, 5), "º")) +
  scale_y_continuous("O3 Wave-1\nPhase",
                     breaks = seq(-pi, pi, by = 30*pi/180),
                     labels = function(x) paste0(round(x*180/pi, 5), "º"))  +
  plot_layout(ncol  = 1) &
  geom_point()  &
  geom_smooth(method = "lm") & 
  panel_background

@@

  # browser()
  stream_gdata %>% 
    periodic(lon = c(0, 360)) %>%
    ggplot(aes(lon, lat)) +

@@

    scale_fill_divergent_discretised("SST (K)",
                                     labels = one_less,
                                     guide = guide_colorsteps_bottom(width = 15,
                                                                     show.limits = FALSE, ticks = TRUE,
                                                                     order = 1)) +
    ggnewscale::new_scale_fill() +
    geom_contour_fill(aes(z = estimate*1e-7, fill = ..level..),
                      breaks = breaks_stream) +
    geom_contour_pval(aes(z = p_val)) +
    # geom_streamline(aes(dx = fx, dy = fy), min.L = 3, L = 10,
    #                 size = 0.1, res = 2, nx = 22, ny = 12, 
    #                 arrow.angle = 14, arrow.length = 0.2) +
    geom_vector(aes(dx = fx, dy = fy), skip = 3, min.mag = 0.002, 
                size = 0.1) +
    scale_mag(max_size = .5, guide = "none") +
    geom_qmap(keep = .015) +
    scale_fill_divergent_discretised(latex2exp::TeX("Streamfunction ($m^2s^{-1}\\times 1e^{-7}$)"), 
                                     labels = one_less,
                                     low = scales::muted("red"),
                                     high = scales::muted("blue"),
                                     guide = guide_colorsteps_bottom(15, show.limits = FALSE,
                                                                     order = 99)) +
    grid_panel(spoke, aes(angle = -angle, var = "SST"), inherit.aes = FALSE) +
    scale_x_continuous(NULL, expand = c(0, 0), 
                       breaks = seq(0, 360, by = 60),
                       labels = LonLabel) +

@@

```{r enso-phase, fig.cap = "(ref:enso-phase-cap)", fig.width = width_column, fig.height=3, include = TRUE}
gdata <- ceof[, eof[[1]]$left, by = .(season)] %>%
  .[cEOF == "cEOF2"] %>% 
  # .[, hgt := Anomaly(hgt)] %>% 
  .[, arg := Arg(hgt)] %>%
  .[, mag := Mod(hgt)] %>%
  .[enso, on = "time"] %>%
  na.omit() %>%
  .[, q := ifelse(mag >= quantile(mag, 0.5), "|cEOF2| > 50%", "|cEOF2| < 50%")]

@@

equation <- latex2exp::TeX(paste0("$ONI = ", 
                                  signif(coef(cos_model)[1], 2), 
                                  "sin(Phase)$"))

@@

  geom_smooth(data = ~qwrap(.x, arg = c(-pi, pi) ~ c(-2*pi, 2*pi)),
              method = "lm", aes(weight = mag),
              formula = y ~ I(sin(x)) -1,
              se = FALSE, color = "black", size = 0.5) +
  geom_point(aes(color = q,
                 shape  = q),
             size = 3) +
  ggrepel::geom_text_repel(data = ~.x[order(-oni)][1:3],
                           aes(label = year(time))) +

>>>>

r2 <- expression(`$r^2$` = paste("", "r", phantom()^{
  paste("2")
}, "", ""))
```

@@

@@

# From latex2exp::TeX('Geopotential height ($m^2s^{-1}$)')
mpg <- expression('Geopotential height ('*m^{2}*s^{phantom() - 1}*')')
```

@@

```{r eof1-regr-gh, fig.cap = "(ref:eof1-regr-gh-cap)", fig.height=1.2*width_column, fig.width=width_column, dependson="plot_regr"}
plot_regr(regr , "cEOF1", c(200, 50), "hgt") +
  scale_fill_divergent(mpg,
                       breaks = AnchorBreaks(0, exclude = 0),
                       guide = guide_colorsteps_bottom(15, even.steps = TRUE), 
                       limits = c(-2000, 2000)) 
```

@@

```{r eof2-regr-gh, fig.cap = "(ref:eof2-regr-gh-cap)", fig.height=1.3*width_column, fig.width=width_column, dependson="plot_regr"}
plot_regr(regr, "cEOF2", c(200, 50), "hgt") +
  scale_fill_divergent(mpg, 
                       breaks = AnchorBreaks(0, 200, exclude = 0),
                       guide = guide_colorsteps_bottom(15, even.steps = TRUE), 
                       limits = c(-1000, 1000))
```

@@

# temp %>% 
#   ggplot(aes(Mod(hgt), amplitude)) +
#   scale_x_continuous("EOF1 Amplitude") +
#   scale_y_continuous("O3 Wave-1\nAmplitude") +
#   
#   temp %>% 
#   copy() %>%
#   na.omit() %>% 
#   qwrap(phase = c(0, 2*pi) ~ c(-pi, pi)) %>% 
#   ggplot(aes(Arg(hgt), phase)) +
#   scale_x_continuous("EOF1 Phase",
#                      breaks = seq(-pi, pi, by = 15*pi/180),
#                      labels = function(x) paste0(round(x*180/pi, 5), "º")) +
#   scale_y_continuous("O3 Wave-1\nPhase",
#                      breaks = seq(-pi, pi, by = 30*pi/180),
#                      labels = function(x) paste0(round(x*180/pi, 5), "º"))  +
#   plot_layout(ncol  = 1) &
#   geom_point()  &
#   geom_smooth(method = "lm") & 
#   panel_background

@@

  #from latex2exp::TeX("Streamfunction ($m^2s^{-1}\\times 1e^{-7}$)")
  psi_expr <- expression('Streamfunction ('*m^{2}*s^{phantom() - 1} %*% 1*e^{phantom() - 7}*')')
  # browser()
  stream_gdata %>% 
    periodic(lon = c(0, 360)) %>%
    ggplot(aes(lon, lat)) +

@@

    scale_fill_divergent_discretised("SST (K)",
                                     labels = one_less,
                                     guide = guide_colorsteps_bottom(width = 15,
                                                                     show.limits = FALSE, ticks = TRUE,
                                                                     order = 1)) +
    ggnewscale::new_scale_fill() +
    geom_contour_fill(aes(z = estimate*1e-7, fill = ..level..),
                      breaks = breaks_stream) +
    geom_contour_pval(aes(z = p_val)) +
    # geom_streamline(aes(dx = fx, dy = fy), min.L = 3, L = 10,
    #                 size = 0.1, res = 2, nx = 22, ny = 12, 
    #                 arrow.angle = 14, arrow.length = 0.2) +
    geom_vector(aes(dx = fx, dy = fy), skip = 3, min.mag = 0.002, 
                size = 0.1) +
    scale_mag(max_size = .5, guide = "none") +
    geom_qmap(keep = .015) +
    scale_fill_divergent_discretised(psi_expr, 
                                     labels = one_less,
                                     low = scales::muted("red"),
                                     high = scales::muted("blue"),
                                     guide = guide_colorsteps_bottom(15, show.limits = FALSE,
                                                                     order = 99)) +
    grid_panel(spoke, aes(angle = -angle, var = "SST"), inherit.aes = FALSE) +
    scale_x_continuous(NULL, expand = c(0, 0), 
                       breaks = seq(0, 360, by = 60),
                       labels = LonLabel) +

@@

```{r enso-phase, fig.cap = "(ref:enso-phase-cap)", fig.width = width_column, fig.height=3}
gdata <- ceof[, eof[[1]]$left, by = .(season)] %>%
  .[cEOF == "cEOF2"] %>% 
  # .[, hgt := Anomaly(hgt)] %>% 
  .[, arg := Arg(hgt)] %>%
  .[, mag := Mod(hgt)] %>%
  .[enso, on = "time"] %>%
  na.omit() %>%
  .[, q := ifelse(mag >= quantile(mag, 0.5), "|cEOF2| > 50%", "|cEOF2| < 50%")]

@@

equation <- bquote(ONI  * {phantom() == phantom()} * .(signif(coef(cos_model)[1], 2)) * sin(Phase) )

@@

  geom_smooth(data = ~qwrap(.x, arg = c(-pi, pi) ~ c(-2*pi, 2*pi)),
              method = "lm", aes(weight = mag),
              formula = y ~ I(sin(x)) -1,
              se = FALSE, color = "black", size = 0.5) +
  geom_point(aes(color = q,
                 shape  = q),
             size = 3) +
  ggrepel::geom_text_repel(data = ~.x[order(-oni)][1:3],
                           aes(label = year(time)), size = 2) +
